NDNS multilevel LCA CW2CB2 or CW2CB3

### 2018-08-28 paper/Chapters/Chapter3.tex
language checked by Hunter

A series of traditional LCA of the responses to carbohydrate intake within 7-time slots of the day was first examined. These initial analyses ignored the clustering of observation days within participants of the survey. \textbf{Table \ref{tab:mixmodels}} shows the latent class solutions for one to five classes (see rows under the Fixed effects model section). The BIC declines with the number of day level classes increases. However, the improvement of BIC dropped to less than 1000 from 3 classes to 4 classes solutions (658.9) and from 4 classes to 5 classes solutions (361.7). Entropy index indicates that the 4 classes model could explain about 51\% percent of the data, while \textit{p} values of Lo-Mendell-Rubun LRT suggest that the more classes we fit, the better model we will have until up to 6 classes (\textit{p} = 0.06 and is not shown in the table). From the parsimony point of view, we extended the model with random effects building on 2 classes, 3 classes, and 4 classes solutions.

The results of the random effect included models are presented in \textbf{Table \ref{tab:mixmodels}} under the Random effects model section. It is obvious that the BIC improves with the addition of the random effects which account for the nested structure of the data. Entropy indicates that 2 classes in the day level and 4 classes in the individual level may be the best solution mathematically. However, after these solutions were checked in more details, the potentially most substantively interpretable model was found to be the 3$\times$3 random effect model, which is the model with 3 latent classes in the day level, and 3 latent classes in the individual level. We must emphasize that different researchers may have made decision slightly different from ours, we have provided other solutions in \textbf{Appendix \ref{AppendixC}} for reference.

@@

	\caption{\label{tab:day-level-features}Means (standard deviations), and counts (\%) of the characteristics of different types of carbohydrate eating days.}\vspace{-0.3cm}
	\centering
	\fontsize{9}{11}\selectfont
	\begin{tabular}[t]{lcccc}
		\hiderowcolors
		\toprule
		& \textbf{\Centerstack{High perc-\\entage carb}} & \textbf{\Centerstack{Low perc-\\entage carb}} & \textbf{\Centerstack{Regular\\meals}} & \textbf{\textit{P} value}\textsuperscript{*}\\
		\midrule
		\showrowcolors
		Counts (\%) & 9667 (39.5) & 5002 (20.4) & 9814 (40.1) & \\
		Country (\%) &  &  &  & < 0.001\\
		\hspace{1em}England & 5627 (58.2) & 2972 (59.4) & 5291 (53.9) & \\
		\hspace{1em}Northern Ireland & 1194 (12.4) & 527 (10.5) & 1400 (14.3) & \\
		\hspace{1em}Scotland & 1527 (15.8) & 813 (16.3) & 1774 (18.1) & \\
		\hspace{1em}Wales & 1318 (13.6) & 690 (13.8) & 1349 (13.7) & \\
		Day of Week (\%) &  &  &  & < 0.001\\
		\hspace{1em}Monday & 1303 (13.5) & 715 (14.3) & 1370 (14.0) & \\
		\hspace{1em}Tuesday & 1266 (13.1) & 674 (13.5) & 1290 (13.1) & \\
		\hspace{1em}Wednesday & 1225 (12.7) & 740 (14.8) & 1233 (12.6) & \\
		\hspace{1em}Thursday & 1272 (13.2) & 752 (15.0) & 1425 (14.5) & \\
		\hspace{1em}Friday & 1458 (15.1) & 797 (15.9) & 1479 (15.1) & \\
		\hspace{1em}Saturday & 1537 (15.9) & 703 (14.1) & 1495 (15.2) & \\
		\hspace{1em}Sunday & 1605 (16.6) & 621 (12.4) & 1522 (15.5) & \\
		\hspace{1em}Weekend, Yes (\%) & 3142 (32.5) & 1324 (26.5) & 3017 (30.7) & < 0.001\\
		Total energy (kJ) & 7539.98 (2875.87) & 7160.22 (2922.15) & 7439.68 (2978.91) & < 0.001\\
		Carbohydrate (g) & 222.79 (89.84) & 209.70 (86.17) & 206.59 (84.42) & < 0.001\\
		Protein (g) & 71.36 (29.79) & 69.55 (30.20) & 73.29 (32.94) & < 0.001\\
		Fat (g) & 65.44 (33.27) & 63.94 (33.76) & 67.24 (34.73) & < 0.001\\
		Alcohol (g) & 11.76 (27.31) & 8.85 (24.25) & 13.80 (33.00) & < 0.001\\
		Total sugars (g) & 98.63 (56.03) & 88.03 (50.50) & 86.39 (50.96) & < 0.001\\
		Starch (g) & 124.07 (55.84) & 121.59 (56.13) & 120.11 (54.62) & < 0.001\\
		Non-milk extrinsic sugar\textsuperscript{\dag} & 59.45 (49.31) & 50.07 (43.41) & 50.41 (44.84) & < 0.001\\
		Fruit (g) & 107.40 (137.97) & 103.15 (129.08) & 92.76 (126.02) & < 0.001\\
		Yellow Red Green Vegetables (g) & 26.52 (46.44) & 26.84 (47.99) & 26.16 (45.99) & 0.681\\
		\bottomrule
		\multicolumn{5}{l}{{\scriptsize \textit{Note:} carb is short for carbohydrate.}}\\
		\multicolumn{5}{l}{{\scriptsize \textsuperscript{*} \textit{P} values were obtained from Pearson $\chi^2$ test for categorical variables, and one-way ANOVA comparing the means}}\\
		\multicolumn{5}{l}{{\scriptsize  in multiple groups for continuous variables;}}\\
		\multicolumn{5}{l}{{\scriptsize \textsuperscript{\dag} Non-milk extrinsic sugar is defined as: additionally added free sugar, such as table sugar, honey, glucose, fructose}}\\ 
		\multicolumn{5}{l}{{\scriptsize and glucose syrups, sugars added to food and sugars in fruit juices.}}\\
	\end{tabular}
\end{table}

@@

With two individual-level latent classes \textbf{(Figure \ref{fig:CB2level2})}, one individual class is comprised of individuals with a relatively slightly higher proportion of having "low carbohydrate day" (22.1\%) compared to the other (17.4\%). This class represents nearly 65\% of the individuals. However, we believe these individual classes are not very distinguishable from each other.

With three individual-level latent classes \textbf{(Figure \ref{fig:level2})}, a low-carbohydrate eaters class, a moderate-carbohydrate eaters class, and a high-carbohydrate eaters class emerge. 43.1\% participants were identified as high-carbohydrate eaters, in these individuals, about 50\% of the days (2 out of 4 days) of their dietary diary could be classified as having high percentage carbohydrate days. Nearly 1 out of 4 days of their dietary diary were either "regular meals day" or "low carbohydrate day". 28.1\% participants fell into the low carbohydrate eaters class on the left-hand side of \textbf{Figure \ref{fig:level2}}, their recordings of food intake showed that in more than 60\% of their days, they were having "regular meals" which was characterised as with the highest amount of fat and alcohol consumptions as already described in \textbf{Table \ref{tab:day-level-features}}. Moderate carbohydrate eaters have comparable proportions (42.0\% vs. 40.0\%) of having high carbohydrate days and regular meals day, 18.0\% of their dietary diary was found to be low carbohydrate days.

@@

Among the three types of carbohydrate eaters, the mean of total energy intake over the 4 days of dietary survey was the highest (7985.8 kJ, 95\%CI: 7283.3, 8146.3) in the low carbohydrate eaters group, and the lowest (7341.8 kJ, 95\%CI: 7172.5, 7511.2) in the moderate eaters group \textbf{(Table \ref{tab:tab-nutri-indi})}. Sources of energy for each type of carbohydrate eaters by the 7-time slots were also different. Low carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-A)} never had carbohydrate contributed more than 50\% of their total energy throughout the day. Energy from fat was the highest for low carbohydrate eaters most of the time during the day (except for the time slot between 10 pm to 6 am next morning). Most impressively, energy contributed by alcohol was always the highest in low carbohydrate eaters, percentages for energy from alcohol for the 7-times slots were 1.4\% (6-9 am), 1.8\% (9-12 noon), 3.1\% (12-2 pm), 7.2\% (2-5 pm), 8.4\% (5-8 pm), 20.0\% (8-10 pm), and 31.2\% (10-6 am), respectively. Contributions from different energy sources were quite similar for moderate and high carbohydrate eaters, but their absolute amount of energy consumption at each time slot was largely different. Moderate carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-B)} were characterised as consuming the lowest energy (296.9 kJ) before 9 am, but having higher energy consumption (967.9 kJ) between 9 am and 12 noon time compared with low and high carbohydrate eaters. Moderate carbohydrate eaters may turn to have later breakfast, later lunch, and probably later dinner as well. They had the highest total energy consumption (599.7 kJ) at night (10 pm - 6 am) across three types of eaters. High carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-C)} consumed the highest total energy (929.0 kJ) during 6 am to 9 am in the morning and the lowest total energy between 10 pm to 6 am (205.5 kJ). Specifically, carbohydrate contributions to total energy intake were 65.3\% (6-9 am), 56.2\% (9-12 noon), 46.2\% (12-2 pm), 50.3\% (2-5 pm), 42.5\% (5-8 pm), 44.1\% (8-10 pm), and 49.9\% (10-6am). We also noticed that high carbohydrate eaters consumed their energy mainly from the following three time slots: 6-9 am, 12-2 pm, and 5-8 pm.

In total, the mean of total carbohydrate intake was 203.8 g, 218.3 g, and 233.4 g for low, moderate, and high carbohydrate eaters, respectively \textbf{(Table \ref{tab:tab-nutri-indi})}. Total energy contributed by carbohydrate was close to 50\% among the high carbohydrate eaters but was only 40.6\% among the low carbohydrate eaters. In terms of the subtypes (components) of the carbohydrate consumed at each time slot, high carbohydrate eaters consumed as much as more than 2 times (compared to low carbohydrate eaters) and nearly 4 times (against moderate carbohydrate eaters) the amount of sugar (37.9g 95\% CI: 36.8, 39.2) and non-milk extrinsic sugar (i.e. free sugar) (11.1g 95\%CI: 10.7, 11.6) between 6-9 am. Moderate carbohydrate eaters had their carbohydrate intake more spread out. They consumed more sugar and starch during 9-12 noon, 2-5 pm, 8-10 pm, and 10-6 am. Low carbohydrate eaters turned to have similar temporal pattern of consuming carbohydrates with the high carbohydrate eaters, but the absolute amount of fibre, sugar, free sugar, and starch were usually lower than that in the high carbohydrate eaters except for time slots of 2-5 pm, and 10-6 am. Strong evidence (\textit{p} < 0.001) suggested that the mean of to-

@@

tal fibre consumption for low, moderate, and high carbohydrate eaters were different: 13.7g (13.4, 14.0), 12.5g (12.1, 12.9), and 14.7g (14.4, 14.9) with all 95\% CI being exclusive to each other. It is also noteworthy that low carbohydrate eaters consumed the highest average amount of protein (79.9 g, 17.2\% of total energy), fat (74.7g, 35.4\% of total energy), and alcohol (20.8 g, 6.8\% of total energy) as we have described for \textbf{Figure \ref{fig:energysourcesCB}}.

The social-demographic characteristics of the UK adults according to their individual level latent class membership are shown in \textbf{Table \ref{tab:Level2tab1}}. Moderate carbohydrate eaters were relatively younger (\textit{p} < 0.001). Gender distribution across the three types of carbohydrate eaters was fairly even (\textit{p} = 0.119). The distribution of the carbohydrate eater types turned out to be changing with the year of survey. Low carbohydrate eaters represented 32.5\% of the population in the first year of the survey but later dropped to lower than 30\% (lowest in the third year, 22.6\%) until the most recent year. The proportion of high carbohydrate eaters increased from 41.2\% to the highest (50.6\%) in the second year, but then started to decline to 38.4\% in the 8th year of survey (\textit{p} = 0.015). There was no evidence of any difference in employment status across three types of carbohydrate eaters. However, strong evidence suggested that high carbohydrate eaters had the highest proportion (61.3\%) of living with partner (\textit{p} < 0.001); moderate carbohydrate eaters had the lowest average income (27180.8 \textsterling/year), highest proportion of non-white population (20.5\%), and lower education level (23.3\% with degree of higher education) compared with either low or high carbohydrate eaters.

Weighted means, percentages of anthropometric measurements, as well as biochemical characteristic profiles according to the latent carbohydrate eater groups are given in \textbf{Table \ref{tab:tab2}}. Low carbohydrate eaters had higher mean BMI (27.8 kg/m\textsuperscript{2}) and larger mean WC (98.9/89.9 cm in men/women) compared with 27.2, 27.3 kg/m\textsuperscript{2}, and 95.9/88.7 (men/women), 98.1/87.2 (men/women) cm in moderate and high carbohydrate eaters. Moderate carbohydrate eaters had the highest prevalence of being a current smoker (27.8\%), shortest time of daily physical activity (geometric mean: 0.87 hours/day), and the lowest prevalence of hypertension (20.2\%).

@@

\textbf{Table \ref{tab:tab1hypetension}} presents the characteristics of men and women participants in the NDNS RP 2008/09-15/16 by hypertension status. The weighted prevalences of hypertension were 30.4\% in men and 27.5\% in women. Among both sexes, there were strong evidence of differences by hypertension status for age, education level, living with a partner or not, smoking status, BMI, abdominal obesity (WC), and prevalence of diabetes (\textit{p} < 0.01). No difference was found among either men or women for ethnicity. Strong evidence of difference was suggested in women for average household income (32741.5 \textsterling/year in non-hypertensive vs. 27862.0 \textsterling/year in hypertensive, \textit{p} < 0.001), and physical activity level (geometric mean: 0.81 hours/day in non-hypertensive compared with 0.53 hours/day in hypertensive, \textit{p} < 0.001) but not in men. Interestingly, in both sexes, hypertensive participants had a higher proportion of being classified as low or high carbohydrate eaters; total energy and carbohydrate intake were higher in people without hypertension (\textit{p} < 0.001).

The sex-specific associations of carbohydrate eating patterns with hypertension (both in total and in participants without diabetes) are shown in \textbf{Table \ref{tab:tab2HYT}}. In the crude models, moderate carbohydrate eaters had statistically significant lower odds of having hypertension than low carbohydrate eaters in both men and women irrespective to diabetes status. Among men, after adjustment for selected confounders, which includes: age, live with partner or not, education level, BMI, smoking status, and total energy intake, the odds ratio (OR) comparing moderate with low carbohydrate eaters was 0.68 (95\% CI: 0.43, 1.07) and remained borderline significant (\textit{p} = 0.093). 95\% CI of the adjusted OR became narrower (OR: 0.64, 95\% CI: 0.41, 1.01, \textit{p} = 0.054) when BMI was replaced with WC in model 2. When diabetic men were excluded in Model 2, the ORs (95\%CI) for moderate and high carbohydrate eaters compared with low carbohydrate eaters were 0.65 (0.41, 1.03) and 0.73 (0.51, 1.06), respectively. The negative associations between moderate carbohydrate eating pattern and hypertension were also observed in women, however, without any statistically significant evidence in the fully adjusted models. High carbohydrate eaters also had lower adjusted odds compared with low carbohydrate eaters, while the 95\% CIs for the adjusted ORs were all wide and included the null value suggesting no evidence of any association in either men or women for high carbohydrate eating pattern and hypertension.

@@

level (\textit{p} = 0.022 for men, < 0.001 for women), average household income (\textit{p} = 0.011 for men, < 0.001 for women) were decreasing with elevated BMI. Living with a partner or not was strongly positively associated with obesity in men but not in women. Men with obesity were also found to have the lowest proportion of never being a smoker (47.9 \%), and the highest proportion of being an ex-smoker (32.9\%). Association between smoking status and obesity in women was only with very weak evidence (\textit{p} = 0.042) but similar pattern as in men was also observed (higher proportion of ex- and current smokers in overweight or obese women). No difference was found for length of physical activity across obesity levels in men, while in women, somewhat weak inverse association (\textit{p}=0.038) was confirmed. Interestingly, predefined carbohydrate eating patterns were seemingly not associated with BMI in men (\textit{p} = 0.072) or in women (\textit{p} = 0.253).

@@

Results of the multivariable linear regression analyses showed inverse associations between latent classes of carbohydrate eating patterns and BMI among men \textbf{(Table \ref{tab:tab2BMI})}. However, the 95\%CI of the regression coefficients were all wide and included the null value 0, indicating no statistically supported evidence for the inverse association. In women, evidence of interaction effect was found in whether live with a partner or not on the association between carbohydrate eating patterns and BMI (\textit{p} for interaction = 0.014 and 0.036 for women in total and without diabetes). For women who were 5.43living with partners, latent classes of carbohydrate eating patterns were negatively associated with BMI. Compared with women eating low carbohydrate food pattern, women having a high carbohydrate eating pattern were averagely associated with 1.76 kg/m\textsuperscript{2} lower BMI after adjustment of age, average household income, education level, smoking status, total energy intake and alcohol consumption. 95\%CI for the adjusted BMI difference was 0.73 to 2.78 kg/m\textsuperscript{2}, \textit{p} = 0.001. After excluding diabetic women, BMI was still 1.62 kg/m\textsuperscript{2} (95\%CI: 0.58, 2.65, \textit{p} = 0.002) smaller in high carbohydrate eaters versus low carbohydrate eaters on average. On the contrary, latent classes of carbohydrate eating patterns were positively associated with BMI in women who were living by themselves, although the regression coefficients were not statistically different from 0.

>>>>

A series of traditional LCA of the responses to carbohydrate intake within the 7 time slots of the day was first examined. These initial analyses ignored the clustering of observation days within participants of the survey. \textbf{Table \ref{tab:mixmodels}} shows the latent class solutions for one to five classes (see rows under the Fixed effects model section). The BIC declines with increases in the number of day level classes increases. However, the improvement of BIC dropped to less than 1000 from 3 class to 4 class solutions (658.9) and from 4 class to 5 class solutions (361.7). Entropy index indicates that the 4 class model could explain about 51\% percent of the data, while \textit{p} values of Lo-Mendell-Rubun LRT suggest that the more classes we fit, the better model we will have until up to 6 classes (\textit{p} = 0.06 and is not shown in the table). For the sake of parsimony, we only extended the model with random effects building on 2 class, 3 class, and 4 class solutions.

The results of the random effects models are presented in \textbf{Table \ref{tab:mixmodels}} under the Random effects model section. It is obvious that the BIC improves with the addition of random effects to account for the nested structure of the data. Entropy indicates that 2 classes in the day level and 4 classes in the individual level may be the best solution mathematically. However, after these solutions were checked in more details, the most substantively interpretable model was found to be the 3$\times$3 random effect model, which is the model with 3 latent classes in the day level and 3 latent classes in the individual level. We must emphasize that different researchers may have made decisions slightly different from ours, we have provided other solutions in \textbf{Appendix \ref{AppendixC}} for reference.

@@

	\caption{\label{tab:day-level-features}Means (standard deviations), and counts (\%) of the characteristics of different types of carbohydrate eating days.}\vspace{-0.3cm}
	\centering
	\fontsize{9}{11}\selectfont
	\begin{tabular}[t]{lcccc}
		\hiderowcolors
		\toprule
		& \textbf{\Centerstack{High perc-\\entage carb}} & \textbf{\Centerstack{Low perc-\\entage carb}} & \textbf{\Centerstack{Regular\\meals}} & \textbf{\textit{P} value}\textsuperscript{*}\\
		\midrule
		\showrowcolors
		Counts (\% in row) & 9667 (39.5) & 5002 (20.4) & 9814 (40.1) & \\
		Country (\%) &  &  &  & < 0.001\\
		\hspace{1em}England & 5627 (58.2) & 2972 (59.4) & 5291 (53.9) & \\
		\hspace{1em}Northern Ireland & 1194 (12.4) & 527 (10.5) & 1400 (14.3) & \\
		\hspace{1em}Scotland & 1527 (15.8) & 813 (16.3) & 1774 (18.1) & \\
		\hspace{1em}Wales & 1318 (13.6) & 690 (13.8) & 1349 (13.7) & \\
		Day of Week (\%) &  &  &  & < 0.001\\
		\hspace{1em}Monday & 1303 (13.5) & 715 (14.3) & 1370 (14.0) & \\
		\hspace{1em}Tuesday & 1266 (13.1) & 674 (13.5) & 1290 (13.1) & \\
		\hspace{1em}Wednesday & 1225 (12.7) & 740 (14.8) & 1233 (12.6) & \\
		\hspace{1em}Thursday & 1272 (13.2) & 752 (15.0) & 1425 (14.5) & \\
		\hspace{1em}Friday & 1458 (15.1) & 797 (15.9) & 1479 (15.1) & \\
		\hspace{1em}Saturday & 1537 (15.9) & 703 (14.1) & 1495 (15.2) & \\
		\hspace{1em}Sunday & 1605 (16.6) & 621 (12.4) & 1522 (15.5) & \\
		\hspace{1em}Weekend, Yes (\%) & 3142 (32.5) & 1324 (26.5) & 3017 (30.7) & < 0.001\\
		Total energy (kJ) & 7539.98 (2875.87) & 7160.22 (2922.15) & 7439.68 (2978.91) & < 0.001\\
		Carbohydrate (g) & 222.79 (89.84) & 209.70 (86.17) & 206.59 (84.42) & < 0.001\\
		Protein (g) & 71.36 (29.79) & 69.55 (30.20) & 73.29 (32.94) & < 0.001\\
		Fat (g) & 65.44 (33.27) & 63.94 (33.76) & 67.24 (34.73) & < 0.001\\
		Alcohol (g) & 11.76 (27.31) & 8.85 (24.25) & 13.80 (33.00) & < 0.001\\
		Total sugars (g) & 98.63 (56.03) & 88.03 (50.50) & 86.39 (50.96) & < 0.001\\
		Starch (g) & 124.07 (55.84) & 121.59 (56.13) & 120.11 (54.62) & < 0.001\\
		Non-milk extrinsic sugar\textsuperscript{\dag} & 59.45 (49.31) & 50.07 (43.41) & 50.41 (44.84) & < 0.001\\
		Fruit (g) & 107.40 (137.97) & 103.15 (129.08) & 92.76 (126.02) & < 0.001\\
		Yellow Red Green Vegetables (g) & 26.52 (46.44) & 26.84 (47.99) & 26.16 (45.99) & 0.681\\
		\bottomrule
		\multicolumn{5}{l}{{\scriptsize \textit{Note:} carb is short for carbohydrate.}}\\
		\multicolumn{5}{l}{{\scriptsize \textsuperscript{*} \textit{P} values were obtained from Pearson $\chi^2$ test for categorical variables, and one-way ANOVA comparing the means}}\\
		\multicolumn{5}{l}{{\scriptsize  in multiple groups for continuous variables;}}\\
		\multicolumn{5}{l}{{\scriptsize \textsuperscript{\dag} Non-milk extrinsic sugar is defined as: additionally added free sugar, such as table sugar, honey, glucose, fructose}}\\ 
		\multicolumn{5}{l}{{\scriptsize and glucose syrups, sugars added to food and sugars in fruit juices.}}\\
	\end{tabular}
\end{table}

@@

With two individual-level latent classes \textbf{(Figure \ref{fig:CB2level2})}, one individual class comprises individuals with a relatively slightly higher proportion of having "low carbohydrate day" (22.1\%) compared to the other (17.4\%). This class represents nearly 65\% of the individuals. However, we believe these individual classes are not very distinguishable from each other.

With three individual-level latent classes \textbf{(Figure \ref{fig:level2})}, a low-carbohydrate eaters class, a moderate-carbohydrate eaters class, and a high-carbohydrate eaters class emerge. 43.1\% participants were identified as high-carbohydrate eaters, in these individuals, about 50\% of the days (2 out of 4 days) of their dietary diary could be classified as having high percentage carbohydrate days. Nearly 1 out of 4 days of their dietary diary were either "regular meals day" or "low carbohydrate day". 28.1\% participants fell into the low carbohydrate eaters class on the left-hand side of \textbf{Figure \ref{fig:level2}}, their recordings of food intake showed that in more than 60\% of their days, they were having "regular meals" which was characterised as with the highest amount of fat and alcohol consumptions as already described in \textbf{Table \ref{tab:day-level-features}}. Moderate carbohydrate eaters have comparable proportions (42.0\% vs. 40.0\%) of having high carbohydrate days and regular meals days; 18.0\% of their dietary diary was found to be low carbohydrate days.

@@

Among the three types of carbohydrate eaters, the mean of total energy intake over the 4 days of dietary survey was the highest (7985.8 kJ, 95\%CI: 7283.3, 8146.3) in the low carbohydrate eaters group, and the lowest (7341.8 kJ, 95\%CI: 7172.5, 7511.2) in the moderate eaters group \textbf{(Table \ref{tab:tab-nutri-indi})}. Sources of energy for each type of carbohydrate eaters by the 7-time slots were also different. Low carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-A)} never had carbohydrate contributed more than 50\% of their total energy throughout the day. Energy from fat was the highest for low carbohydrate eaters most of the time during the day (except for the time slot between 10 pm to 6 am next morning). Most impressively, energy contributed by alcohol was always the highest in low carbohydrate eaters, percentages for energy from alcohol for the 7-times slots were 1.4\% (6-9 am), 1.8\% (9-12 noon), 3.1\% (12-2 pm), 7.2\% (2-5 pm), 8.4\% (5-8 pm), 20.0\% (8-10 pm), and 31.2\% (10-6 am), respectively. Contributions from different energy sources were quite similar for moderate and high carbohydrate eaters, but their absolute amount of energy consumption at each time slot was largely different. Moderate carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-B)} were characterised as consuming the lowest energy (296.9 kJ) before 9 am but having higher energy consumption (967.9 kJ) between 9 am and 12 noon time compared with low and high carbohydrate eaters. Moderate carbohydrate eaters may tend to have later breakfast, later lunch, and probably later dinner as well. They had the highest total energy consumption (599.7 kJ) at night (10 pm - 6 am) across all three types of eaters. High carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-C)} consumed the highest total energy (929.0 kJ) during 6 am to 9 am in the morning and the lowest total energy between 10 pm to 6 am (205.5 kJ). Specifically, carbohydrate contributions to total energy intake were 65.3\% (6-9 am), 56.2\% (9-12 noon), 46.2\% (12-2 pm), 50.3\% (2-5 pm), 42.5\% (5-8 pm), 44.1\% (8-10 pm), and 49.9\% (10-6 am). We also noticed that high carbohydrate eaters consumed their energy mainly from the following three time slots: 6-9 am, 12-2 pm, and 5-8 pm.

In total, the mean of total carbohydrate intake was 203.8 g, 218.3 g, and 233.4 g for low, moderate, and high carbohydrate eaters, respectively \textbf{(Table \ref{tab:tab-nutri-indi})}. Total energy contributed by carbohydrate was close to 50\% among the high carbohydrate eaters but was only 40.6\% among the low carbohydrate eaters. In terms of the subtypes (components) of the carbohydrate consumed at each time slot, high carbohydrate eaters consumed more than 2 times (compared to low carbohydrate eaters) and nearly 4 times (against moderate carbohydrate eaters) the amount of sugar (37.9g 95\% CI: 36.8, 39.2) and non-milk extrinsic sugar (i.e. free sugar, 11.1g 95\%CI: 10.7, 11.6) between 6-9 am. Moderate carbohydrate eaters had carbohydrate intakes that were more spread out throughout the day. They consumed more sugar and starch during 9-12 noon, 2-5 pm, 8-10 pm, and 10-6 am. Low carbohydrate eaters turned to have similar temporal pattern of consuming carbohydrates with the high carbohydrate eaters, but the absolute amount of fibre, sugar, free sugar, and starch consumed were usually lower than that in the high carbohydrate eaters except for time slots of 2-5 pm, and 10-6 am. Strong evidence (\textit{p} < 0.001) suggested that the

@@

mean of total fibre consumption for low, moderate, and high carbohydrate eaters were different: 13.7g (95\%CI: 13.4, 14.0), 12.5g (95\%CI: 12.1, 12.9), and 14.7g (95\%CI: 14.4, 14.9) with all 95\% CI being exclusive to each other. It is also noteworthy that low carbohydrate eaters consumed the highest average amount of protein (79.9 g, 17.2\% of total energy), fat (74.7g, 35.4\% of total energy), and alcohol (20.8 g, 6.8\% of total energy) as we have described for \textbf{Figure \ref{fig:energysourcesCB}}.

The social-demographic characteristics of the UK adults according to their individual level latent class membership are shown in \textbf{Table \ref{tab:Level2tab1}}. Moderate carbohydrate eaters were relatively younger (\textit{p} < 0.001). Gender distribution across the three types of carbohydrate eaters was fairly even (\textit{p} = 0.119). The distribution of the carbohydrate eater types appears to be changing with the year of survey. Low carbohydrate eaters represented 32.5\% of the population in the first year of the survey but later dropped to lower than 30\% (lowest in the third year, 22.6\%) until the most recent year. The proportion of high carbohydrate eaters increased from 41.2\% to the highest (50.6\%) in the second year, but then started to decline to 38.4\% in the 8th year of survey (\textit{p} = 0.015). There was no evidence of any difference in employment status across three types of carbohydrate eaters. However, strong evidence suggested that high carbohydrate eaters had the highest proportion (61.3\%) of living with partner (\textit{p} < 0.001); moderate carbohydrate eaters had the lowest average income (27180.8 \textsterling/year), the highest proportion of non-white population (20.5\%), and a lower education level (23.3\% with degree of higher education) compared with either low or high carbohydrate eaters.

Weighted means, percentages of anthropometric measurements, as well as biochemical characteristic profiles according to the latent carbohydrate eater groups are given in \textbf{Table \ref{tab:tab2}}. Low carbohydrate eaters had higher mean BMI (27.8 kg/m\textsuperscript{2}) and larger mean WC (98.9/89.9 cm in men/women) compared with 27.2, 27.3 kg/m\textsuperscript{2}, and 95.9/88.7 (men/women), 98.1/87.2 (men/women) cm in moderate and high carbohydrate eaters. Moderate carbohydrate eaters had the highest prevalence of being a current smoker (27.8\%), shortest time of daily physical activity (geometric mean: 0.87 hours/day), and lowest prevalence of hypertension (20.2\%).

@@

\textbf{Table \ref{tab:tab1hypetension}} presents the characteristics of men and women participants in the NDNS RP 2008/09-15/16 by hypertension status. The weighted prevalences of hypertension were 30.4\% in men and 27.5\% in women. Among both sexes, there was strong evidence of differences by hypertension status for age, education level, living with a partner or not, smoking status, BMI, abdominal obesity (WC), and prevalence of diabetes (\textit{p} < 0.01). No difference was found among either men or women for ethnicity. Strong evidence of difference was suggested in women for average household income (32741.5 \textsterling/year in non-hypertensive vs. 27862.0 \textsterling/year in hypertensive, \textit{p} < 0.001), and physical activity level (geometric mean: 0.81 hours/day in non-hypertensive compared with 0.53 hours/day in hypertensive, \textit{p} < 0.001) but not in men. Interestingly, in both sexes, hypertensive participants had a higher proportion of being classified as low or high carbohydrate eaters; total energy and carbohydrate intake were higher in people without hypertension (\textit{p} < 0.001).

The sex-specific associations of carbohydrate eating patterns with hypertension (both in total and in participants without diabetes) are shown in \textbf{Table \ref{tab:tab2HYT}}. In the crude models, moderate carbohydrate eaters had statistically significant lower odds of having hypertension than low carbohydrate eaters in both men and women irrespective to diabetes status. Among men, after adjustment for selected confounders, which includes: age, live with partner or not, education level, BMI, smoking status, and total energy intake, the odds ratio (OR) comparing moderate with low carbohydrate eaters was 0.68 (95\% CI: 0.43, 1.07) and remained borderline significant (\textit{p} = 0.093). 95\% CI of the adjusted OR became narrower (OR: 0.64, 95\% CI: 0.41, 1.01, \textit{p} = 0.054) when BMI was replaced with WC in model 2. When diabetic men were excluded in Model 2, the odds for moderate and high carbohydrate eaters compared with low carbohydrate eaters were 35\% (OR: 0.65, 95\%CI: 0.41, 1.03) and 27\% (OR:0.73, 95\%CI: 0.51, 1.06) lower, respectively. The negative associations between moderate carbohydrate eating pattern and hypertension were also observed in women, however, without any statistically significant evidence in the fully adjusted models. High carbohydrate eaters also had lower adjusted odds compared with low carbohydrate eaters, while the 95\% CIs for the adjusted ORs were all wide and included the null value suggesting no evidence of any association in either men or women for high carbohydrate eating pattern and hypertension.

@@

level (\textit{p} = 0.022 for men, < 0.001 for women), and average household income (\textit{p} = 0.011 for men, < 0.001 for women) decreased with increases in BMI. Living with a partner or not was strongly positively associated with obesity for men but not for women. Men with obesity were also found to have the lowest proportion of never being a smoker (47.9\%) and the highest proportion of being an ex-smoker (32.9\%). Association between smoking status and obesity in women had only very weak evidence (\textit{p} = 0.042), but a similar pattern as in men was observed (higher proportion of ex- and current smokers in overweight or obese women). No difference was found for length of physical activity across obesity levels in men, while in women, somewhat weak inverse association (\textit{p}=0.038) was confirmed. Interestingly, predefined carbohydrate eating patterns lacked evidence for an association with BMI in men (\textit{p} = 0.072) or in women (\textit{p} = 0.253).

@@

Results of the multivariable linear regression analyses showed inverse associations between latent classes of carbohydrate eating patterns and BMI among men \textbf{(Table \ref{tab:tab2BMI})}. However, the 95\%CI of the regression coefficients were all wide and included the null value 0, indicating no statistically supported evidence for the inverse association. In women, evidence of an interaction effect was found between living with a partner or not and carbohydrate eating patterns for the outcome of BMI (\textit{p} for interaction = 0.014 and 0.036 for women in total and without diabetes). For women who were living with partners, latent classes of carbohydrate eating patterns were negatively associated with BMI. Compared with women eating a low carbohydrate food pattern, women having a high carbohydrate eating pattern were associated with an average 1.76 kg/m\textsuperscript{2} lower BMI after adjustment of age, average household income, education level, smoking status, total energy intake and alcohol consumption. 95\%CI for the adjusted BMI difference was 0.73 to 2.78 kg/m\textsuperscript{2}, \textit{p} = 0.001. After excluding diabetic women, BMI was still 1.62 kg/m\textsuperscript{2} (95\%CI: 0.58, 2.65, \textit{p} = 0.002) smaller in high carbohydrate eaters versus low carbohydrate eaters on average. On the contrary, latent classes of carbohydrate eating patterns were positively associated with BMI in women who were living by themselves, although there was no evidence that regression coefficients different from 0 (\textit{p} > 0.05).
